Gravitational wave sources in our Galactic backyard: LISA predictions for BH and NS binaries

### 2021-06-18 paper/tex/sections/comparisons.tex
update comparisons

In Figure~\ref{fig:previous_studies}, we compare similar previous studies that have investigated the population of stellar mass BHBHs, BHNSs and NSNSs that are detectable with LISA. The table details the expected detection rates predicted by each paper as well as their assumptions regarding their Milky Way galaxy model, binary population synthesis simulation and LISA mission specifications. This table not include the number papers on the LISA WDWD population as we do not make predictions for these DCOs.

\citet{Belczynski+2010} built upon this work by using a different population synthesis code with two model variations and a multi-component model for the Milky Way. They found a similar number of BHBH and BHNS detection to \citet{Nelemans+2001}, but a lower number of NSNS detections. The low total detection rate for all DCOs in this paper is unsurprising given the relatively high SNR threshold of 10 and short mission length of 1 year. The reduced mission length means that the source signal has much less time to accumulate, whilst many fewer WDWDs can be resolved in this time, leading to a weaker signal and an increased Galactic confusion noise relative to our work.

The discrepancy between their results and ours could be a result of different implicit assumptions in their population synthesis code, however we expect that BSE should be similar to COMPAS since it provided the basis of the COMPAS code. Another major difference is that, unlike this work, they assume that all binaries are circular for the purpose of detection in LISA, which could result in a lower number of detections by missing eccentric binaries that appear as weaker signals when assumed to be circular. We also improve upon this work by using a larger number of metallicity bins since a low number of metallicity bins can produce artificial features in the mass distribution of DCOs and possibly affect the detection rate \todo{needs citation, perhaps @Stephen knows?}.

\citet{Lau+2020} investigated the number of Galactic NSNS binaries that could be detected by LISA. Their study uses the same population synthesis code, COMPAS, as this work, though an earlier version. Despite this, their study found a much larger number of detections. They make several different physical assumptions in their population synthesis, using the \citet{Fryer+2012} \textit{rapid} remnant mass prescription, limiting the maximum neutron star mass to $2 \unit{M_{\odot}}$ and not implementing PISN. However, we note that none of these assumptions significantly affect the NSNS LISA detection rate (see bottom panel of Fig.\,\ref{fig:detection_rates}, models \modRapid{}, \modNSLow{} and \modNoPISN{}).

Finally, \citet{Breivik+2020} introduced the population synthesis code COSMIC and presented detections for many different DCO types in LISA using this code. \todo{what I write here will depend how much the numbers change in upcoming run. Long story short is that they only have two metallicity bins and use the optimistic CE model.}.

% find that LISA will detect 93 BHBH, 33 BHNS and 8 NSNS binaries in the Milky Way over a 4 year mission. Although these are significantly different from our fiducial results, we make many different physical assumptions, the most notable being that \citet{Breivik+2020} assumes the optimistic CE scenario. Thus it would be more prudent to compare to our results from model \modOpt{} in which we find 70, 38 and 15 detections respectively. The discrepancy in the NSNS rate is likely due to the fact that we assume that case BB mass transfer onto a NS is always stable but \citet{Breivik+2020} does not and this would significantly decrease our rate (see bottom panel of Fig.\,\ref{fig:detection_rates}, model \modCaseBB{}) . The remaining differences are likely due to using a different population synthesis code (COSMIC) and using a different model for the Milky Way, particularly the assumptions of two fixed metallicities.

@@

>>>>

In Figure~\ref{fig:previous_studies}, we compare similar previous studies that have investigated the population of stellar mass BHBHs, BHNSs and NSNSs that are detectable with LISA. The table details the expected detection rates predicted by each paper as well as their assumptions regarding their Milky Way galaxy model, binary population synthesis simulation and LISA mission specifications. This table not include the number papers on the LISA WDWD population as we do not make predictions for these DCOs. We additionally do not include papers that don't use population synthesis or do not simulate sources in the Galactic plane \citep[e.g.][]{Andrews+2020, Kremer+2018} such that we only compare with papers with similar methodology and focus.

@@

\citet{Belczynski+2010} built upon this work by using a different population synthesis code with two model variations and a multi-component model for the Milky Way. They found a similar number of BHBH and BHNS detection to \citet{Nelemans+2001}, but a lower number of NSNS detections. The low total detection rate for all DCOs in this paper is unsurprising given the relatively high SNR threshold of 10 and short mission length of 1 year. The reduced mission length means that the source signal has much less time to accumulate, whilst also fewer WDWDs can be resolved in this time, leading to a weaker signal and an increased Galactic confusion noise relative to our work.

The discrepancy between their results and ours could be a result of different implicit assumptions in their population synthesis code, however we expect that BSE should be similar to COMPAS since it provided the basis of the COMPAS code. Another major difference is that, unlike this work, they assume that all binaries are circular for the purpose of detection in LISA, which could result in a lower number of detections by missing eccentric binaries that appear as weaker signals when assumed to be circular. We also improve upon this work by using a larger number of metallicity bins since a low number of metallicity bins can produce artificial features in the mass distribution of DCOs and possibly affect the detection rate (Kummer et al. (in prep)).

\citet{Lau+2020} investigated the number of Galactic NSNS binaries that could be detected by LISA. Their study uses the same population synthesis code, COMPAS, as this work, though an earlier version. Despite this, their study found a much larger number of detections. They make several different physical assumptions in their population synthesis, using the \citet{Fryer+2012} \textit{rapid} remnant mass prescription, limiting the maximum neutron star mass to $2 \unit{M_{\odot}}$ and not implementing PISN. However, we note that none of these assumptions strongly affect the NSNS LISA detection rate (see bottom panel of Fig.\,\ref{fig:detection_rates}, models \modRapid{}, \modNSLow{} and \modNoPISN{}).

Finally, \citet{Breivik+2020} introduced the population synthesis code COSMIC and presented detections for many different DCO types in LISA using this code. They find that LISA will detect 93 BHBH, 33 BHNS and 8 NSNS binaries in the Milky Way over a 4 year mission. They make many different physical assumptions, the most notable being that \citet{Breivik+2020} assumes the optimistic CE scenario. Thus it would be more prudent to compare to our results from model \modOpt{} in which we find 118, 52 and 7 detections respectively. Thus the NSNS rate is consistent but we find higher rates for the BHBHs and BHNSs. These differences are likely due to using a different population synthesis code (COSMIC) and using a different model for the Milky Way, particularly the assumptions of two fixed metallicities.
